### 2025-03-26 ms/ms_old.tex
Initial commit for tree-allo

\begin{document}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, breakable, frame hidden, enhanced, sharp corners, boxrule=0pt, interior hidden]}{\end{tcolorbox}}\fi

@@

Identifying the most appropriate mathematical representation of
allometric relationships is a subject of ongoing debate. While numerous
papers have used various functional forms for allometric equations
(e.g., \protect\hyperlink{ref-Thomas1996}{Thomas 1996};
\protect\hyperlink{ref-Feldpausch2011}{Feldpausch et al. 2011};
\protect\hyperlink{ref-MartinezCano2019}{Martínez Cano et al. 2019};
\protect\hyperlink{ref-Song2024}{Song et al. 2024}), power-law functions
remain the most common choice for describing the allometric scaling of
tree size variables, including height and crown size with DBH (e.g.,
\protect\hyperlink{ref-Sileshi2023}{Sileshi et al. 2023};
\protect\hyperlink{ref-Jucker2024}{Jucker et al. 2024}). The
relationships among tree size variables often involve complex
interactions influenced by various factors, such as species diversity,
environmental conditions, and individual variability (e.g.,
\protect\hyperlink{ref-Poorter2012a}{Poorter et al. 2012a};
\protect\hyperlink{ref-Laurans2024}{Laurans et al. 2024}). Consequently,
power-law models often fail to accurately capture the allometric
relationships across the full range of tree sizes, leading to
underestimation of dimensions for smaller trees and overestimation for
larger trees (e.g., \protect\hyperlink{ref-Fayolle2016}{Fayolle et al.
2016}; \protect\hyperlink{ref-Song2024}{Song et al. 2024}). This
highlights the need to adopt alternative functional forms---such as
saturating relationships---that can more accurately reflect the
complexity of allometric scaling in vegetation models (e.g.,
\protect\hyperlink{ref-Goodman2014}{Goodman et al. 2014};
\protect\hyperlink{ref-Song2024}{Song et al. 2024}).

Earth observation (EO) technologies, including remote sensing and Light
Detection and Ranging (LiDAR), have revolutionized the monitoring of
forests (e.g., \protect\hyperlink{ref-Jucker2017}{Jucker et al. 2017})
by enabling the identification and measurement of individual tree
heights and crown dimensions using airborne imagery (e.g.,
\protect\hyperlink{ref-Shendryk2016}{Shendryk et al. 2016};
\protect\hyperlink{ref-Jucker2017}{Jucker et al. 2017}). However,
because existing allometric models still rely on stem diameter as a
crucial input for biomass estimation, this reliance presents a challenge
for effectively integrating EO technologies into AGB estimation
(\protect\hyperlink{ref-Shendryk2016}{Shendryk et al. 2016};
\protect\hyperlink{ref-Jucker2017}{Jucker et al. 2017}). While
ground-based LiDAR can measure DBH directly and is much more efficient
than traditional ground surveys
(\protect\hyperlink{ref-Tatsumi2023}{Tatsumi et al. 2023}), it is still
less common and convenient than aerial LiDAR. Alternative approaches
focusing on tree height and crown dimensions as model predictors are
also necessary for remote sensing applications
(\protect\hyperlink{ref-Jucker2017}{Jucker et al. 2017}). One potential
solution is to estimate DBH from tree height and crown size and then use
established biomass equations, which could be an effective strategy for
integrating airborne imagery into carbon monitoring programs
(\protect\hyperlink{ref-Jucker2017}{Jucker et al. 2017}).

Accurately representing tree allometry relationships also requires
considering underlying ecological constraints, particularly the
interplay of plant functional traits and their trade-offs influence
individual tree architecture and the composition of entire forest
communities (e.g., \protect\hyperlink{ref-Iida2012}{Iida et al. 2012};
\protect\hyperlink{ref-LoubotaPanzou2018}{Loubota Panzou et al. 2018};
\protect\hyperlink{ref-Yang2023}{Yang and Swenson 2023}). Under
closed-canopy conditions, strong light competition makes the balance
between vertical and horizontal space use crucial for species
coexistence (e.g., \protect\hyperlink{ref-Aiba2009}{Aiba and Nakashizuka
2009}; \protect\hyperlink{ref-Iida2012}{Iida et al. 2012}). Among
functional traits, wood density is closely linked to both plant life
history strategies (\protect\hyperlink{ref-Fajardo2022a}{Fajardo 2022})
and light demand (e.g., \protect\hyperlink{ref-King2006}{King et al.
2006}; \protect\hyperlink{ref-Aiba2009}{Aiba and Nakashizuka 2009};
\protect\hyperlink{ref-Poorter2012a}{Poorter et al. 2012a}). Several
studies suggest that wood density explains a trade-off between rapid
stem expansion and broader crowns, shaping mechanical constraints on
species with diverse architecture and light capture strategies
(\protect\hyperlink{ref-Aiba2009}{Aiba and Nakashizuka 2009};
\protect\hyperlink{ref-Iida2012}{Iida et al. 2012};
\protect\hyperlink{ref-Yang2023}{Yang and Swenson 2023}). However, these
types of studies have examined this relationship post hoc, relating
species-specific allometric parameters to wood density and risking Type
I or Type II errors if the hierarchical data structure is overlooked
(\protect\hyperlink{ref-Gelman2006}{Gelman and Hill 2006}). One
exception is Martínez Cano et al.
(\protect\hyperlink{ref-MartinezCano2019}{2019}), who incorporated wood
density into a hierarchical model of tropical trees and concluded it
offered minimal explanatory power for interspecific variation.
Therefore, the role of wood density in global allometric scaling remains
insufficiently explored.

In this study, we present a robust Bayesian hierarchical approach to
delineate allometric relationships involving DBH, tree height (H), crown
radius (CR), and their interspecific variations. Using large datasets
from the Tallo and TRY databases
(\protect\hyperlink{ref-Kattge2020}{Kattge et al. 2020};
\protect\hyperlink{ref-Jucker2022}{Jucker et al. 2022}), which span a
diverse array of tree forms across global forest types and climatic
conditions, we overcome the limitations of previous studies that often
relied on site-specific or regional observations with limited samples
and species. By integrating species-level wood density (WD) into a
multilevel model, our methodology avoids the post-hoc correlations used
in previous studies, offering a more universal and scalable model,
effectively capturing variability from the species to the community
level. Our main objective is to develop models informed by tree
allometric variables (H, CR, DBH, WD) to capture interspecific variation
in allometric scaling globally and offer an improved framework for
estimating above ground biomass. Specifically, we ask the following
questions: (1) What is the best function to estimate tree architecture
from DBH and \emph{vice versa}? (2) How does wood density shape
interspecific variation in these allometric scaling relationships? (3)
How does the choice of allometric relationships affect subsequent AGB
estimation?

@@

Our analysis combined tree allometry data from Tallo, which includes
standardized measurements of DBH, height, and crown radius for 5,163
species across various biomes (\protect\hyperlink{ref-Jucker2022}{Jucker
et al. 2022}, https://doi.org/10.5281/zenodo.6637599) along with wood
density data for 10,995 species from TRY version 6
(\protect\hyperlink{ref-Kattge2020}{Kattge et al. 2020},
https://www.try-db.org/TryWeb/Home.php) (see Supplementary Table S1 for
further details). We thoroughly processed the Tallo dataset by removing
998 outliers following Jucker et al.
(\protect\hyperlink{ref-Jucker2022}{2022}). To ensure consistency, we
selected only species with standardized wood density values in g
cm\textsuperscript{-3} from TRY and removed any records with missing
data. Subsequently, we created an overlapping dataset between Tallo and
TRY, resulting in a combined dataset encompassing allometric and wood
density information for 355,881 individuals across 2,961 species.

@@

Once the best-performing models for each allometric relationship were
identified, we tested whether adding wood density as a species-level
predictor would further improve the predictive performance of those
models and also examined its influence on tree architecture diversity.
Species-specific parameters (\(\beta_j\)) were modeled as a function of
wood density (\(u_j\)): \begin{equation}\protect\hypertarget{eq-wood}{}{
\beta_{j} \sim N(\gamma_{0} + \gamma_{1} u_j, \sigma_{\beta}),
}\label{eq-wood}\end{equation}

where \(\gamma_{0}\) and \(\gamma_{1}\) are the species-level intercept
and slope, respectively, and \(\sigma_{\beta}\) is the standard
deviation of the species-specific parameters. To facilitate
interpretation and visualization, species were then classified into
``light'' and ``dense'' wood groups based on the 10th and 90th
percentiles of wood density distribution in the observed dataset,
respectively. This classification allowed us to illustrate the
relationships between wood density and the parameters of the
best-performing models, enabling an evaluation of how wood density
contributes to variation in tree architecture. Additionally, we explored
a model that accounts for covariance between species-specific parameters
using a multivariate normal distribution. However, due to the complexity
of this approach, the model did not converge for the nonlinear models,
and thus, we did not include it in the final analysis.

@@

For angiosperm species, we used the widely applied equation from Chave
et al. (\protect\hyperlink{ref-Chave2014}{2014}) to estimate aboveground
biomass (AGB; kg of dry mass):

@@

The optimal model to predict tree height was a Weibull function for both
angiosperms (ang) and gymnosperms (gym) (Table 1). Across the global
dataset, the best model to predict H from DBH of each clade was:
\begin{equation}\protect\hypertarget{eq-h_ang}{}{
\mathrm{H_{ang}} =50.0
\left\{ 1 - \exp\left( - 0.0476
\mathrm{DBH}^{0.629}
\right) \right\}
}\label{eq-h_ang}\end{equation}

@@

The second-best model, which used the gMM function, showed poorer
predictive performance in angiosperms (\(\Delta\)LOOIC = 101.01) but
showed competitive predictive performance in gymnosperms
(\(\Delta\)LOOIC = 4.87). While power-law models (Eq.~\ref{eq-pl}) can
exhibit saturating behavior when 0 \textless{} \emph{b} \textless{} 1,
they performed the worst overall compared to other saturating models
(Eq.~\ref{eq-gMM}; Eq.~\ref{eq-wb}). Incorporating wood density did not
significantly improve predictive performance (Table 1).

\textbf{Table 1}: Summary of the results of the model selection
procedure. Models were ranked using the Leave-One-Out Information
Criterion (LOOIC) to assess their predictive performance. For each
model, we report the Expected Log Predictive Density (ELPD), the LOOIC,
and the \(R^2\) values. The difference in LOOIC relative to the
best-performing model, \(\Delta\)LOOIC, is also included, with a
\(\Delta\)LOOIC greater than 4 suggests a meaningful difference in
predictive performance (\protect\hyperlink{ref-Cochrane2023}{Cochrane et
al. 2023}; \protect\hyperlink{ref-Sivula2023}{Sivula et al. 2023}).
Lower \(\Delta\)LOOIC values indicate models that perform closer to the
best model in terms of predictive accuracy, while higher ELPD values
reflect better model fit. Higher \(R^2\) values reflect better
explanatory power, indicating how well the model explains the variance
in the dependent variable.

@@

\caption{\label{fig-compare}Allometric relationships between DBH and
tree height or crown radius in angiosperms (a, c) and gymnosperms (b,
d), based on three different models: power-law (orange), generalized
Michaelis-Menten (green), and Weibull (blue). A color gradient from
black (low density) to yellow (high density) indicates the density of
overlapping data points, calculated via 2D kernel density estimation on
a 300\,\(\times\)\,300 grid. See Supplementary Figure S1 for the
log-scale version of these plots.}

@@

The power-law function (0 \textless{} \emph{b} \textless{} 1) provided
the best predictive performance for modeling the relationship between
crown radius and DBH in angiosperms. Whereas for gymnosperms, the gMM
function offered the best performance for this allometric scaling (Table
1). Across the global dataset, the best model for predicting CR from DBH
of each clade was: \begin{equation}\protect\hypertarget{eq-cr_ang}{}{
\mathrm{CR_{ang}} = \mathrm{0.349}
\, \mathrm{DBH}^{\mathrm{0.626}}
}\label{eq-cr_ang}\end{equation}

@@

For angiosperms, the saturating functions (Eq.~\ref{eq-gMM};
Eq.~\ref{eq-wb}) performed worse than the power-law model, with
\(\Delta\)LOOIC values of 124.57 for gMM and 126.35 for the Weibull
function. In gymnosperms, the Weibull function showed a competitive
performance compared to the gMM function, while the power-law function
(Eq.~\ref{eq-pl}) performed the worst, with a \(\Delta\)LOOIC of 162.72.
Incorporating wood density did not significantly improve predictive
performance (Table 1).

\textbf{Table 2}: Posterior estimates of the parameters for the best
hierarchical models selected for tree height, crown radius, and DBH
allometries in angiosperms and gymnosperms. Each set includes posterior
estimates both without and with the inclusion of wood density, reported
as median and 95\% Bayesian credible intervals for the community-level
parameters of each allometric function. For tree height, the Weibull
model was identified as the best for both clades. Crown radius was best
described by a power-law model in angiosperms and a gMM model for
gymnosperms. The best model for DBH allometry in both clades, which
incorporates both tree height and crown radius information (\(CR,H\)
model).

@@

The \(CR,H\) model, which treated CR and H as separate factors,
exhibited the best predictive performance among all tested models (Table
1). Across the global dataset, the most effective model for predicting
DBH from H and CR for each clade was:
\begin{equation}\protect\hypertarget{eq-dbh_ang}{}{
\mathrm{DBH_{ang}} = \mathrm{0.948}
\mathrm{CR}^{\text{0.374}}
\mathrm{H}^{\text{0.590}}
}\label{eq-dbh_ang}\end{equation}

@@

\caption{\label{fig-wood2}Effect of wood density on tree height
allometry in angiosperms (a) and on crown radius allometry in
angiosperms (b) and gymnosperms (c). Curves represent the allometric
scaling for species with light (orange) and dense (brown) wood density,
corresponding to the 10\% and 90\% quantiles of wood density
distribution in the observed dataset, respectively. The shaded areas
illustrate the 95\% Bayesian credible intervals. Only statistically
significant associations from Fig.~\ref{fig-wood1} are plotted.}

@@

AGB estimates based on the Weibull model (\(\text{AGB}_\text{H-WB}\))
showed the highest predictive accuracy based both on the lowest root
mean square error (RMSE) of 0.279 Mg and the smallest bias of 2.04\%
(Fig.~\ref{fig-agb} a). Estimates based on the power-law model
(\(\text{AGB}_\text{H-PL}\)) also showed competitive performance but had
higher RMSE and bias compared to the Weibull-based approach, with RMSE
of 0.28 Mg and bias of 2.29\% (Fig.~\ref{fig-agb} b). In contrast,
models that used estimated DBH performed substantially worse
(Fig.~\ref{fig-agb} c-f).

@@

\caption{\label{fig-agb}Comparison of individual tree above-ground
biomass (AGB) estimated using different approaches with reference AGB
(\(\text{AGB}_\text{ref}\)) derived from observed DBH and tree height in
angiosperms. (a, b) AGB estimates based on observed DBH and tree height
predicted by the Weibull function (\(\text{AGB}_\text{H-WB}\)) or the
power-law function (\(\text{AGB}_\text{H-PL}\)). (c--f) AGB estimates
based on observed tree height and DBH predicted from crown radius and
tree height as separate factors
(\(\text{AGB}_{\text{DBH} \sim \text{CR,H}}\)), as a compound term
(\(\text{AGB}_{\text{DBH} \sim \text{CR × H}}\)), from tree height alone
(\(\text{AGB}_{\text{DBH} \sim \text{H}}\)), or from crown radius alone
(\(\text{AGB}_{\text{DBH} \sim \text{CR}}\)). Dashed lines represent the
1:1 relationships, and the solid lines show linear regression fits to
the data, highlighting how predictive accuracy varies with tree size.
RMSE (root mean square error) and bias values for each model are
reported.}

@@

Our result supported a saturating H-DBH relationship in both angiosperm
and gymnosperm species, consistent with theoretical expectations
(\protect\hyperlink{ref-Falster2003}{Falster and Westoby 2003};
\protect\hyperlink{ref-Niklas2007}{Niklas 2007}) and empirical
observations (e.g., \protect\hyperlink{ref-Feldpausch2011}{Feldpausch et
al. 2011}; \protect\hyperlink{ref-MartinezCano2019}{Martínez Cano et al.
2019}; \protect\hyperlink{ref-Song2024}{Song et al. 2024}). A recent
study using Tallo suggested the Weibull function as the best model for
six plant functional types (PFTs) (\protect\hyperlink{ref-Song2024}{Song
et al. 2024}). While PFTs-based approaches can illuminate large-scale
patterns, they may obscure species-specific variation in allometric
scaling, and locally abundant species may bias the overall estimates.
Our best predictive model effectively captures species-specific patterns
(Supplementary Table S6) despite underestimating the height of
individuals with DBH \textgreater{} 30 cm due to substantial
interspecific variation in H--DBH allometric scaling
(Fig.~\ref{fig-compare}). This highlights the advantages of a multilevel
modeling approach in addressing systematic differences across species.
Additionally, the metabolic scaling theory (MST) posits logarithmic
scaling relationships among different axes of tree size, with an
exponent close to 2/3 (\(H \propto DBH^{2/3}\),
\(CR \propto DBH^{2/3}\)), based on assumptions regarding mechanical and
hydraulic constraints to plant growth
(\protect\hyperlink{ref-West1999}{West et al. 1999}). However, our
findings revealed that the power-law model exponent of angiosperms
(\emph{b} = 0.538) was significantly smaller, while the actual exponent
in gymnosperms (\emph{b} = 0.704) was larger than the theoretical value
of 2/3 (Supplementary Table S3). This divergence highlights the
importance of separately modeling the allometric relationships of each
functional group to accurately capture their distinct scaling patterns.
Moreover, the power-law function performed worse than the saturating
functions (Table 1), indicating that the actual H-DBH relationship does
not follow the simple scaling laws proposed by MST.

>>>>

\begin{document}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, boxrule=0pt, enhanced, frame hidden, interior hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi

@@

Identifying the most appropriate mathematical representation of
allometric relationships is a subject of ongoing debate. While numerous
papers have used various functional forms for allometric equations
(e.g., \protect\hyperlink{ref-Thomas1996}{Thomas 1996};
\protect\hyperlink{ref-Feldpausch2011}{Feldpausch et al. 2011};
\protect\hyperlink{ref-MartinezCano2019}{Martínez Cano et al. 2019};
\protect\hyperlink{ref-Song2024}{Song et al. 2024}), power-law functions
remain the most common choice for describing the allometric scaling of
tree size variables, including height and crown size with DBH (e.g.,
\protect\hyperlink{ref-Sileshi2023}{Sileshi et al. 2023};
\protect\hyperlink{ref-Jucker2024}{Jucker et al. 2024}). The
relationships among tree size variables often involve complex
interactions influenced by various factors, such as species diversity,
environmental conditions, and individual variability
(\protect\hyperlink{ref-Laurans2024}{Laurans et al. 2024}; e.g.,
\protect\hyperlink{ref-Poorter2012}{\textbf{Poorter2012?}}).
Consequently, power-law models often fail to accurately capture the
allometric relationships across the full range of tree sizes, leading to
underestimation of dimensions for smaller trees and underestimation for
larger trees (e.g., \protect\hyperlink{ref-Fayolle2016}{Fayolle et al.
2016}; \protect\hyperlink{ref-Song2024}{Song et al. 2024}). This
highlights the need to adopt alternative functional forms---such as
saturating relationships---that can more accurately reflect the
complexity of allometric scaling in vegetation models (e.g.,
\protect\hyperlink{ref-Goodman2014}{Goodman et al. 2014};
\protect\hyperlink{ref-Song2024}{Song et al. 2024}).

Earth observation (EO) technologies, including remote sensing and Light
Detection and Ranging (LiDAR), have revolutionized monitoring forests
(e.g., \protect\hyperlink{ref-Jucker2017}{Jucker et al. 2017}) by
enabling the identification and measurement of individual tree heights
and crown dimensions using airborne imagery (e.g.,
\protect\hyperlink{ref-Shendryk2016}{Shendryk et al. 2016};
\protect\hyperlink{ref-Jucker2017}{Jucker et al. 2017}). Because
existing allometric models still depend on stem diameter as a crucial
input for biomass estimation, it also presents the challenge of
determining the most effective way to utilize EO technologies for AGB
estimation (\protect\hyperlink{ref-Shendryk2016}{Shendryk et al. 2016};
\protect\hyperlink{ref-Jucker2017}{Jucker et al. 2017}). While
ground-based LiDAR can measure DBH directly and is much more efficient
than traditional ground surveys
(\protect\hyperlink{ref-Tatsumi2023}{Tatsumi et al. 2023}), it is still
less common and convenient than aerial LiDAR. Alternative approaches
focusing on tree height and crown dimensions as model predictors are
also necessary for remote sensing applications
(\protect\hyperlink{ref-Jucker2017}{Jucker et al. 2017}). One potential
solution is to estimate DBH from tree height and crown size and then use
established biomass equations, which could be an effective strategy for
integrating airborne imagery into carbon monitoring programs
(\protect\hyperlink{ref-Jucker2017}{Jucker et al. 2017}).

Accurately representing tree allometry relationships also requires
considering underlying ecological constraints, particularly the
interplay of plant functional traits and their trade-offs influences
individual tree architecture and the composition of entire forest
communities (e.g., \protect\hyperlink{ref-Iida2012}{Iida et al. 2012};
\protect\hyperlink{ref-LoubotaPanzou2018}{Loubota Panzou et al. 2018};
\protect\hyperlink{ref-Yang2023}{Yang and Swenson 2023}). Under
closed-canopy conditions, strong light competition makes the balance
between vertical and horizontal space use crucial for species
coexistence (e.g., \protect\hyperlink{ref-Aiba2009}{Aiba and Nakashizuka
2009}; \protect\hyperlink{ref-Iida2012}{Iida et al. 2012}). Among
functional traits, wood density is closely linked to both plant life
history strategies
(\protect\hyperlink{ref-Fajardo2022}{\textbf{Fajardo2022?}}) and light
demand (e.g., \protect\hyperlink{ref-King2006}{King et al. 2006};
\protect\hyperlink{ref-Aiba2009}{Aiba and Nakashizuka 2009};
\protect\hyperlink{ref-Poorter2012}{\textbf{Poorter2012?}}). Several
studies suggest that wood density explains a trade-off between rapid
stem expansion and broader crowns, shaping mechanical constraints on
species with diverse architecture and light capture strategies
(\protect\hyperlink{ref-Aiba2009}{Aiba and Nakashizuka 2009};
\protect\hyperlink{ref-Iida2012}{Iida et al. 2012};
\protect\hyperlink{ref-Yang2023}{Yang and Swenson 2023}). However, these
types of studies have examined this relationship post hoc, relating
species-specific allometric parameters to wood density and risking Type
I or Type II errors if the hierarchical data structure is overlooked
(\protect\hyperlink{ref-Gelman2006}{Gelman and Hill 2006}). One
exception is Martínez Cano et al.
(\protect\hyperlink{ref-MartinezCano2019}{2019}), who incorporated wood
density into a hierarchical model of tropical trees and concluded it
offered minimal explanatory power for interspecific variation.
Therefore, the role of wood density in global allometric scaling remains
insufficiently explored.

In this study, we present a robust Bayesian hierarchical approach to
delineate allometric relationships involving DBH, tree height (H), crown
radius (CR), and their interspecific variations. Using large datasets
from the Tallo and TRY databases
(\protect\hyperlink{ref-Kattge2020}{Kattge et al. 2020};
\protect\hyperlink{ref-Jucker2022}{Jucker et al. 2022}), which span a
diverse array of tree forms across global forest types and climatic
conditions, we overcome the limitations of previous studies that often
relied on site-specific or regional observations with limited samples
and species. By integrating species-level wood density into a multilevel
model, our methodology avoids the post-hoc correlations used in previous
studies, offering a more universal and scalable model, effectively
capturing variability from the species to the community level.
Specifically, we ask the following questions: (1) What is the best
function to estimate tree architecture from DBH and \emph{vice versa}?
(2) How does wood density shape interspecific variation in these
allometric scaling relationships? (3) How does the choice of allometric
relationships affect subsequent AGB estimation?

@@

Our analysis combined tree allometry data from Tallo, which includes
standardized measurements of DBH, height, crown radius for 5,163 species
across various biomes (\protect\hyperlink{ref-Jucker2022}{Jucker et al.
2022}, https://doi.org/10.5281/zenodo.6637599) along with wood density
data for 10,995 species from TRY version 6
(\protect\hyperlink{ref-Kattge2020}{Kattge et al. 2020},
https://www.try-db.org/TryWeb/Home.php) (see Supplementary Table S1 for
further details). We thoroughly processed the Tallo dataset by removing
998 outliers following Jucker et al.
(\protect\hyperlink{ref-Jucker2022}{2022}). To ensure consistency, we
selected only species with standardized wood density values in g
cm\textsuperscript{-3} from TRY and removed any records with missing
data. Subsequently, we created an overlapping dataset between Tallo and
TRY, resulting in a combined dataset encompassing allometric and wood
density information for 355,881 individuals across 2,961 species.

@@

Once the best-performing models were identified for each allometric
relationship, we then tested whether adding a species-level predictor
(wood density) would further improve that model's predictive
performance. Species-specific parameters (\(\beta_j\)) were modeled as a
function of wood density (\(u_j\)):
\begin{equation}\protect\hypertarget{eq-wood}{}{
\beta_{j} \sim N(\gamma_{0} + \gamma_{1} u_j, \sigma_{\beta}),
}\label{eq-wood}\end{equation}

where \(\gamma_{0}\) and \(\gamma_{1}\) are the species-level intercept
and slope, respectively, and \(\sigma_{\beta}\) is the standard
deviation of the species-specific parameters. Additionally, we explored
a model that accounts for covariance between species-specific parameters
using a multivariate normal distribution. However, due to the complexity
of this approach, the model did not converge for the nonlinear models,
and thus, we did not include it in the final analysis.

@@

For angiosperm species, we estimated aboveground biomass (AGB; kg of dry
mass) using the equation from Chave et al.
(\protect\hyperlink{ref-Chave2014}{2014}):

@@

The optimal model to predict tree height was a Weibull function for both
angiosperms (ang) and gymnosperms (gym) (Table 1). Across the global
dataset, the best model to predict H from DBH of each division was:
\begin{equation}\protect\hypertarget{eq-h_ang}{}{
\mathrm{H_{ang}} =50.0
\left\{ 1 - \exp\left( - 0.0476
\mathrm{DBH}^{0.629}
\right) \right\}
}\label{eq-h_ang}\end{equation}

@@

The second-best model, which used the gMM function, showed poorer
predictive performance in angiosperms (\(\Delta\) LOOIC = 101.01) but
showed competitive predictive performance in gymnosperms (\(\Delta\)
LOOIC = 4.87). While power-law models (Eq.~\ref{eq-pl}) can exhibit
saturating behavior when 0 \textless{} \emph{b} \textless{} 1, they
performed the worst overall compared to other saturating models
(Eq.~\ref{eq-gMM}; Eq.~\ref{eq-wb}). Incorporating wood density did not
significantly improve predictive performance (Table 1).

\textbf{Table 1}: Summary of the results of the model selection
procedure. Models were ranked using the Leave-One-Out Information
Criterion (LOOIC) to assess their predictive performance. For each
model, we report the Expected Log Predictive Density (ELPD), the LOOIC,
and the \(R^2\) values. The difference in LOOIC relative to the
best-performing model, \(\Delta\)\_LOOIC, is also included. Lower
\(\Delta\)\_LOOIC values indicate models that perform closer to the best
model in terms of predictive accuracy, while higher ELPD values reflect
better model fit. Higher \(R^2\) values reflect better explanatory
power, indicating how well the model explains the variance in the
dependent variable.

@@

\caption{\label{fig-compare}Allometric relationships between tree height
and crown radius with DBH of angiosperm (a, c) and gymnosperm (b, d)
based on three different models: power-law (orange), generalized
Michaelis-Menten (green), and Weibull (blue). The density of overlapping
points is shown using a color gradient ranging from black (low density)
to yellow (high density). Point density was calculated using a 2D kernel
density estimation on a 300 \(\times\) 300 grid, estimating the
concentration of data points in the space of DBH and tree height or
crown radius (see Supplementary Figure S1 for the log-scale version).}

@@

The power-law function (0 \textless{} \emph{b} \textless{} 1) provided
the best predictive performance for modeling the relationship between
crown radius and DBH in angiosperms. Whereas for gymnosperms, the gMM
function offered the best performance for this allometric scaling (Table
1). Across the global dataset, the best model for predicting CR from DBH
of each division was: \begin{equation}\protect\hypertarget{eq-cr_ang}{}{
\mathrm{CR_{ang}} = \mathrm{0.349}
\, \mathrm{DBH}^{\mathrm{0.626}}
}\label{eq-cr_ang}\end{equation}

@@

For angiosperms, the saturating functions (Eq.~\ref{eq-gMM};
Eq.~\ref{eq-wb}) performed worse than the power-law model, with
\(\Delta\) LOOIC values of 124.57 for gMM and 126.35 for the Weibull
function. In gymnosperms, the Weibull function showed a competitive
performance compared to the gMM function, while the power-law function
(Eq.~\ref{eq-pl}) performed the worst, with a \(\Delta\) LOOIC of
162.72. Incorporating wood density did not significantly improve
predictive performance (Table 1).

\textbf{Table 2}: Posterior estimates of the parameters for the best
hierarchical models selected for tree height, crown radius, and DBH
allometries. Each set includes the best models for angiosperms and
gymnosperms, with posterior estimates reported both without and with the
inclusion of wood density. Table entries represent the median and 95\%
Bayesian credible intervals for the community-level parameters of each
allometric function. For tree height, the Weibull model was identified
as the best for both angiosperm and gymnosperm groups. Crown radius was
best described by a power-law model for angiosperms and a gMM model for
gymnosperms. The best model for DBH allometry in both divisions
incorporated both tree height and crown radius information (\(CR,H\)
model).

@@

The \(CR,H\) model, which treated CR and H as separate factors,
exhibited the best predictive performance among all tested models (Table
1). Across the global dataset, the most effective model for predicting
DBH from H and CR for each division was:
\begin{equation}\protect\hypertarget{eq-dbh_ang}{}{
\mathrm{DBH_{ang}} = \mathrm{0.948}
\mathrm{CR}^{\text{0.374}}
\mathrm{H}^{\text{0.590}}
}\label{eq-dbh_ang}\end{equation}

@@

\caption{\label{fig-wood2}Effect of wood density on tree height
allometry in angiosperms (a) and on crown radius allometry in
angiosperms (b) and gymnosperms (c). Curves represent the allometric
scaling for species with light (orange) and dense (brown) wood density,
corresponding to the 10\% and 90\% quantiles of wood density
distribution in the observed dataset, respectively. The shaded areas
illustrate the 95\% CIs. Only statistically significant associations
from Fig.~\ref{fig-wood1} are plotted.}

@@

AGB estimates based on the Weibull model (\(\text{AGB}_\text{H-WB}\))
showed the highest predictive accuracy based both on the lowest root
mean square error (RMSE) of 0.279 Mg and the smallest bias of 2.043\%
(Fig.~\ref{fig-agb} a). Estimates based on the power-law model
(\(\text{AGB}_\text{H-PL}\)) also showed competitive performance but had
higher RMSE and bias compared to the Weibull-based approach, with RMSE
of 0.28 Mg and bias of 2.29\% (Fig.~\ref{fig-agb} b). In contrast,
models that used estimated DBH performed substantially worse
(Fig.~\ref{fig-agb} c-f).

@@

\caption{\label{fig-agb}Comparison of individual tree above-ground
biomass estimated from different approaches with reference AGB
(\(\text{AGB}_\text{ref}\)) derived from observed DBH and tree height in
angiosperms. (a, b) AGB estimates based on observed DBH and tree height
predicted by the Weibull function (\(\text{AGB}_\text{H-WB}\)) or the
power-law function (\(\text{AGB}_\text{H-PL}\)). (c--f) AGB estimates
based on observed tree height and DBH predicted from crown radius and
tree height as separate factors
(\(\text{AGB}_{\text{DBH} \sim \text{CR,H}}\)), as a compound term
(\(\text{AGB}_{\text{DBH} \sim \text{CR × H}}\)), from tree height alone
(\(\text{AGB}_{\text{DBH} \sim \text{H}}\)), or from crown radius alone
(\(\text{AGB}_{\text{DBH} \sim \text{CR}}\)). Dashed lines represent the
1:1 relationships, and the solid lines are linear regression fits to the
data points to highlight how predictive accuracy varies with tree size.
RMSE (root mean square error) and bias values for each model are
reported.}

@@

Our result supported a saturating H-DBH relationship in both angiosperm
and gymnosperm species, consistent with theoretical expectations
(\protect\hyperlink{ref-Falster2003}{Falster and Westoby 2003};
\protect\hyperlink{ref-Niklas2007}{Niklas 2007}) and empirical
observations (e.g., \protect\hyperlink{ref-Feldpausch2011}{Feldpausch et
al. 2011}; \protect\hyperlink{ref-MartinezCano2019}{Martínez Cano et al.
2019}; \protect\hyperlink{ref-Song2024}{Song et al. 2024}). A recent
study using Tallo suggested the Weibull function as the best model for
six plant functional types (PFTs) (\protect\hyperlink{ref-Song2024}{Song
et al. 2024}). While PFTs-based approaches can illuminate large-scale
patterns, they may obscure species-specific variation in allometric
scaling, and locally abundant species may bias the overall estimates.
Our best predictive model effectively captures species-specific patterns
(Supplementary Table S6) despite underestimating the height of
individuals with DBH \textgreater{} 30 cm due to substantial
interspecific variation in H--DBH allometric scaling
(Fig.~\ref{fig-compare}). This highlights the advantages of a multilevel
modeling approach in addressing systematic differences across species.
Additionally, the metabolic scaling theory (MST) posits logarithmic
scaling relationships among different axes of tree size, with an
exponent close to 2/3 (\(H \propto DBH^{2/3}\),
\(CR \propto DBH^{2/3}\)), based on assumptions regarding mechanical and
hydraulic constraints to plant growth
(\protect\hyperlink{ref-West1999}{West et al. 1999}). However, our
findings revealed that the power-law model exponent of angiosperms
(\emph{b} = 0.54) was significantly smaller, while the actual exponent
in gymnosperms (\emph{b} = 0.7) was larger than the theoretical value of
2/3 (Supplementary Table S3). This divergence highlights the importance
of separately modeling the allometric relationships of each functional
group to accurately capture their distinct scaling patterns. Moreover,
the power-law function performed worse than the saturating functions
(Table 1), indicating that the actual H-DBH relationship does not follow
the simple scaling laws proposed by MST.
